---
title: "Interpretation of Data II"
author: "A00275664"
date: "2023-11-19"
output: pdf_document
---

# RMarkdown Guide 

## RMarkdown GP Applications

```{r include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```
RMarkdown can be used by a GP in a variety of ways to optimise and enhance a 
GP's efficiency in day-to-day tasks
Some ways that RMarkdown can be applied are;

  * create reports for team-members
  
  * provide calculations for tax returns
  
  * provide an audit trail for reviews and audits
  
  * communicate knowledge easily with non-technical staff members

### Creating Reports with RStudio

RMarkdown files are clear and straightforward to read.
For example;

The average speed of a car (miles per hour) from the 1920s was
```{r Mean example, echo=FALSE}
data<- cars
mean(cars$speed)
```
### Visualisation Features

Charts and graphs can be produced from data to communicate a clearer picture of 
company performance. RMarkdown utilises the package **'ggplot2'** for data
visualisation.
```{r scatter_chunk, echo=TRUE, fig.height=4}
ggplot(ToothGrowth, aes(x=len,
                        y=dose,
                        color= supp))+
  geom_point() + labs(title="Tooth Length by Dosage",
                      x = "Dose (mg/day)",
                      y = "Tooth Length")
```


This can be shown a variety of ways useful for a business, such as a *bar chart* 
```{r bar_chunk, echo=TRUE, fig.height=3.5}
ggplot(ToothGrowth, aes(x = dose, fill = supp)) +
  geom_bar(position = "dodge") +
  labs(title = "Tooth Growth by Dose and Supplement",
       x = "Dose (mg/day)",
       y = "Frequency")
```
\hfill\break
\hfill\break
Or a *histogram*
```{r hist_chunk, echo=TRUE, fig.height=3.5}
ggplot(ToothGrowth, aes(len))+
  geom_histogram(fill="purple", colour = "black", binwidth = 1)
```

### Summary Statistics

RMarkdown and Rstudio can provide a quick analysis of your data
```{r}
summary(airquality)
```
This can also be narrowed down to a single variable
```{r}
summary(airquality$Ozone)
```

Creating an example dataframe
```{r}
Patients<-data.frame(
  Pat_Id = 1:10,
  Age = c(18,19,20,22,30,40,50,65,70,71),
  Weight = c(65, 70, 80, 75, 85, 90, 120, 80, 95, 100)
  )
```

The packages **'dplyr'** can provide functions such as

  * mutate()  - create new columns based on a function using variables within the dataframe
  * select()  - select variables and create new datasets containing them
  * filter()  - select certain rows that meet certain criteria
  * summarise() - provides summary statistics for a dataframe or variable 
  * arrange() - organise rows in a dataframe

Using the *filter()* function to find all patients weighing above 85 kilograms from the Patients dataframe
```{r}
Patients1<-filter(Patients, Weight > 85)

```
The *mutate()* function can create a new variable from existing information in the dataframe
```{r}
#Asssume height is equal for all
Height<-1.70
Patients2<- Patients %>%
  mutate(BMI = Weight / (Height*2))
```
RStudio allows you to return your output with the *print()* function
```{r}
print(Patients2)
```
\hfill\break
It is also possible to load your own data into RMarkdown using *read.csv()*  and *read.excel()* functions,
depending on the data file type
```{r}
data <- read.csv("C:/Users/erink/Downloads/inpatientCharges.csv")
names(data)
```
Here is an example of the *summarise()* function finding the average patient discharges in the USA
```{r}
average_Total.Discharges <- data %>%
                summarise(avg_Total.Discharges = mean(Total.Discharges))
print(average_Total.Discharges)
```
Lastly using the **Tidyr** package, we can create tidy data using 

  * gather()  - removes specified columns and adds named columns

  * pivot_longer() - reshapes data from a wider format to a longer format.
      
  * spreading()  - lengthens the observations across multiple rows.
      
  * pivot_wider()  - reshapes data from a longer format to a wider format.
      
  * seperating() - splits columns into multiples columns.
      
  * uniting() - combines multiple columns into a single column.

An example using *unite()* to combine the insurance provider's two address columns in the inpatientCharges dataset
```{r}
data1<- unite(data, "Provider_Address",
              Provider.Street.Address, Provider.City, 
              sep =", ")
```
\hfill\break
\hfill\break
\hfill\break



# References

```
E. Green, ‘Answer to “Change R Markdown plot width”’, Stack Overflow. 
Accessed: Nov. 17, 2023. [Online]. Available: https://stackoverflow.com/a/23755105/22916988

Ssayols, ‘Answer to “Line breaks in R Markdown text (not code blocks)”’, 
Stack Overflow. Accessed: Nov. 17, 2023. [Online]. 
Available: https://stackoverflow.com/a/43113246/22916988

‘Hospital Charges for Inpatients’. Accessed: Nov. 17, 2023. [Online]. 
Available: https://www.kaggle.com/datasets/speedoheck/inpatient-hospital-charges

file:///C:/Users/erink/Downloads/rmarkdown%20(1).pdf [R Markdown Cheat Sheet]
```
